A hierarchical approach for modelling X-ray beamlines. Application to a coherent beamline.

### 2019-04-03 paper-hierarchical.tex
Updates from Overleaf

Regarding computational effort, simulations concerning ray tracing (Section~\ref{level1}) and coherent optics (Sections~\ref{wofry} and \ref{srw_se}) can be done interactively in a normal laptop. However, for those concerning partial coherence, very long calculations are needed. For example, Monte Carlo calculations in Section~\ref{srw_me} for the multi electron case \inred{HOW MANY ELECTRONS?} spent a few hours on a 56-core computer cluster. For performing the COMSYL coherent mode decomposition the calculation time depends very much on the dimensions on the problem. For EBS the wavefronts were sampled using 1007x335 pixels, and 1103 modes were computed representing 98\% of the spectral density. This represents 5 MB per wavefront and a matrix of 1695 GB to diagonalize. It took less that four hours in the a cluster of 28 cores. The calculation is more expensive if one wants to decompose a source that is pretty incoherent like our ESRF-High-$\beta$. This is a limit case, it was used wavefields of XxX pixels to accommodate the large spatial extension of the source. The calculation took XX days in a large XX cluster. 
%To give an idea of the real cost, a similar work send to the AWS cluster costed about 2000 USD. 
This was done as a test case, to demonstrate the feasibility of the calculations. It is not recommended to do similar calculations that imply a the cost because a source that is quite incoherent was treated with partial coherence methods. In Appendix B we present for the ESRF High$\beta$ similar results than those in Section~\ref{comsyl}.

@@

We have obtained quantitative values of flux, transmission, size of the focused beam, and coherent fraction for ID16A, a modern beamline with already exceptional performances in terms of nanofucusing and coherence. We predicted advances the improvements in these parameters when the EBS will start in 2020. We obtained these parameters using a hierarchical flowchart with, applying methods with increasing complexity and requiring higher computational effort.

The analytical estimation of the main parameters is mandatory first step in the beamline conception and analysis. We applied concepts based on optical magnification and numerical aperture acceptance and calculated approximately the coherent fraction at the source. These values help to perform sanity check of the computer simulations. A ray tracing simulation is inexpensive and provides important information. The comparison of the CF with the beamline transmitivity would give an idea of the beam coherence, reflected in the number modes present in the final beam. For high coherence, like for ID16A, the wave optics calculations are necessary. If the final beam has a very high coherence, like in our case, coherent optics (1D simplifications, and single electron calculations) may give good results. Full calculations based on coherent mode decomposition Were introduced here. It is a new way of quantitatively assess the quality of the partial coherent beam via the calculation of the coherent modes of the source, propagation of the modes and recalculating the coherent mode expansion on the final beam from which one obtains the CF, a parameter that valutated the coherence quality.

@@

\section{Metrology of the KB mirrors}

>>>>

Regarding computational effort, simulations concerning ray tracing (Section~\ref{level1}) and coherent optics (Sections~\ref{wofry} and \ref{srw_se}) can be done interactively in a normal laptop. However, for those concerning partial coherence, very long calculations are needed. For example, Monte Carlo calculations in Section~\ref{srw_me} for the multi electron case \inred{HOW MANY ELECTRONS?} spent a few hours on a 56-core computer cluster. For performing the COMSYL coherent mode decomposition the calculation time depends very much on the dimensions on the problem. For EBS the wavefronts were sampled using 1007x335 pixels, and 1103 modes were computed representing 98\% of the spectral density. This represents 5 MB per wavefront and a matrix of 1695 GB to diagonalize. It took less that four hours in the a cluster of 28 cores. The calculation is more expensive if one wants to decompose a source that is pretty incoherent like our ESRF-High-$\beta$. This is a limit case, it was used wavefields of 10075x201 pixels to accommodate the large spatial extension of the source. We obtained 4016 modes totalizing only XX\% of the spectral density.  The calculation took more than 4 days in a 84 cores cluster. 
%To give an idea of the real cost, a similar work send to the AWS cluster costed about 2000 USD. 
This was done as a test case, to demonstrate the feasibility of the calculations. It is not recommended to do similar calculations that imply a the cost because a source that is quite incoherent was treated with partial coherence methods. In Appendix B we present for the ESRF High$\beta$ similar results than those in Section~\ref{comsyl}.

@@

We have obtained quantitative values of flux, transmission, size of the focused beam, and coherent fraction for ID16A, a modern beamline with already exceptional performances in terms of nanofucusing and coherence. We predicted advances the improvements in these parameters when the EBS will start in 2020. We obtained these parameters using a hierarchical flowchart applying methods with increasing complexity and requiring higher computational effort.

The analytical estimation of the main parameters is mandatory first step in the beamline conception and analysis. We applied concepts based on optical magnification and numerical aperture acceptance and calculated approximately the coherent fraction at the source. These values help to perform sanity check of the computer simulations. A ray tracing simulation is inexpensive and provides important information. The comparison of the CF with the beamline transmitivity would give an idea of the beam coherence, reflected in the number modes present in the final beam. For high coherence, like for ID16A, the wave optics calculations are necessary. If the final beam has a very high coherence coherent optics (1D simplifications, and single electron calculations) may give good results. Full calculations based on coherent mode decomposition Were introduced here. It is a new way of quantitatively assess the quality of the partial coherent beam via the calculation of the coherent modes of the source, propagation of the modes and recalculation the coherent mode expansion on the final beam. The CF or occupations of the first mode is the parameter used to measure the coherence quality.

@@

\section{Metrology of the KB mirrors}
The residual height error of the multilayer monochromator and the KB mirrors, after removing the main ellitical shape, are shown in Fig.~\ref{fog:metrology}. These profiles are used along the text in ray tracing and wave optics simulations.

\begin{figure}\label{fig:metrology}
\includegraphics[width=0.8\textwidth]{GRAPHICS/ID16A_ML_W-B4C_120Strp.png}
\includegraphics[width=0.8\textwidth]{GRAPHICS/ID16A_KB1_VF.png}
\includegraphics[width=0.8\textwidth]{GRAPHICS/ID16A_KB1_HF.png}
\caption{Error height profiles used for the simulations for a) the ML monochromator, b) the vertical focusing mirror os the KB, and c) T=the horozontal focusing mirror of the KB.}
\end{figure}
